---
title: "COVID-19 Classification"
output:
  html_document:
    df_print: paged
  html_notebook: default
---

# Loading libraries
```{r}
library(dplyr)
library(caret)
library(RANN)
library(rpart)
library(rpart.plot)
library(caTools)
library(ggplot2)
```


# Loading the dataset
```{r}
covid <- read.csv("~/Downloads/Covid Dataset.csv")
```

# EDA
```{r}
head(covid)
str(covid)
summary(covid)
colnames(covid)
sum(is.na(covid)) # no missing values
```


# Initial visualizations
```{r}
ggplot(covid, aes(x = COVID.19, fill = Breathing.Problem)) + geom_bar(position = "dodge") + labs(title = "Relationship Between Breathing Problem and COVID-19 Status", x = "COVID-19 Status", y = "Count") + theme_minimal()
ggplot(covid, aes(x = COVID.19, fill = Contact.with.COVID.Patient)) + geom_bar(position = "dodge") + labs(title = "Contact with COVID Patient and COVID-19 Status", x = "COVID-19 Status", y = "Count") + theme_minimal()
ggplot(covid, aes(x = COVID.19, fill = Visited.Public.Exposed.Places)) + geom_bar(position = "dodge") + labs(title = "Visited Public Exposed Places and COVID-19 Status", x = "COVID-19 Status", y = "Count") + theme_minimal()
ggplot(covid, aes(x = COVID.19, fill = Wearing.Masks)) + geom_bar(position = "dodge") + labs(title = "Wearing Masks and COVID-19 Status", x = "COVID-19 Status", y = "Count") + theme_minimal()
ggplot(covid, aes(x = COVID.19, fill = Fever)) + geom_bar(position = "dodge") + labs(title = "Relationship Between Fever and COVID-19 Status", x = "COVID-19 Status", y = "Count") + theme_minimal()
ggplot(covid, aes(x = COVID.19, fill = Headache)) + geom_histogram(stat = "count", position = "dodge") + labs(title = "Distribution of Headache Across COVID-19 Status", x = "COVID-19 Status", y = "Count") + theme_minimal()
ggplot(covid, aes(x=COVID.19, fill=Asthma)) + geom_bar(position = "dodge") + labs(title = "Relationship between Asthma and COVID-19 Status", x = "COVID-19 Status", y = "Count") + theme_minimal()

asthma_pie <- covid %>% count(Asthma) %>% mutate(percentage = n / sum(n) * 100)
ggplot(asthma_pie, aes(x = "", y = percentage, fill = Asthma)) + geom_col() + coord_polar(theta = "y") + labs(title = "Distribution of Asthma in the Dataset", fill = "Asthma") + theme_void()

sanitization_pie <- covid %>% count(COVID.19, Sanitization.from.Market) %>% group_by(COVID.19) %>% mutate(percentage = n / sum(n) * 100)
ggplot(sanitization_pie, aes(x = "", y = percentage, fill = Sanitization.from.Market)) + geom_col() + coord_polar(theta = "y") + facet_wrap(~ COVID.19) + labs(title = "Sanitization Practices by COVID-19 Status", fill = "Sanitization from Market") + theme_void()

ggplot(covid, aes(x=COVID.19)) + geom_bar() + labs(title="Distribution of COVID-19 In This Dataset", x="COVID-19 Status", y="Count") + theme_minimal()

```


# Re-balancing the data by under sampling the majority class, i.e. reducing the number of Yes's to match the number of No's
```{r}
covid_no <- covid[covid$COVID.19 == "No", ]
covid_yes <- covid[covid$COVID.19 == "Yes", ]
covid_balanced <- rbind(covid_no, covid_yes[sample(1:nrow(covid_yes), nrow(covid_no)), ])

ggplot(covid_balanced, aes(x=COVID.19)) + geom_bar() + labs(title="Distribution of COVID-19 In This Dataset", x="COVID-19 Status", y="Count") + theme_minimal()

```


# Building the model
```{r}
covid_balanced$COVID.19 <- as.factor(covid_balanced$COVID.19)

set.seed(123)
sample = sample.split(covid_balanced$COVID.19, SplitRatio = 0.80)
train = subset(covid_balanced, sample=TRUE)
test = subset(covid_balanced, sample=FALSE)
```

# Building decision tree
```{r}
options(repr.plot.width = 10, repr.plot.height = 10)
covid_tree <- rpart(COVID.19 ~., cp=0.001, maxdepth=5, data=covid_balanced)

```


# Visualizing the model
```{r}
prp(covid_tree, space=4, split.cex=1.2, nn.border.col=0)
```

# Analyzing model performance
```{r}
covid_predictions <- predict(covid_tree, test, type="class")
confusionMatrix(covid_predictions, test$COVID.19)
```


# Extracting feature importance from the decision tree
```{r}
importance <- as.data.frame(varImp(covid_tree))
importance$Feature <- rownames(importance)
rownames(importance) <- NULL
colnames(importance)[1] <- "Importance"
```

# Visualizing feature importance
```{r}
ggplot(importance, aes(x = reorder(Feature, Importance), y = Importance)) + geom_bar(stat = "identity", fill = "steelblue") + coord_flip() + labs(title = "Feature Importance in Predicting COVID-19", x = "Features", y = "Importance") + theme_minimal()
```








